\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-10-08 fully-abstract-comp/popl2019/paper.tex
Address reviewer D comments

%The main contributions of this paper are (1) \stktokens{} a calling convention that uses linear capabilities to provably secure LSE and WBCF, and (2) a novel proof technique that we call fully-abstract overlay semantics which we apply to prove the properties of \stktokens{}.
In this paper, we make the following contributions:
% Presents the first formal model of a capability machine with linear capabilities (Section 2)
% Presents a calling convention called \stktokens{} that provably guarantee local state encapsulation and well-bracketed control flow (Section 3)
% Presents a novel proof technique that we call fully-abstract overlay semantics. (Section 4-5)
\begin{itemize}
\item We present \trgcm{}: A formalization of a simple CHERI-like capability machine with linear capabilities (Section~\ref{sec:cap-mach-w-seal-and-lin}).
\item We present a new calling convention \stktokens{} that provably guarantees LSE and WBCF on \trgcm{} (Section~\ref{sec:stktokens-explained}).
\item We present a new way to formalize these guarantees based on a novel technique called \textit{fully-abstract overlay semantics} and we prove the results. This includes:
  \begin{itemize}
  \item defining \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF (Section~\ref{sec:form-secur-with})
  \item proving full-abstraction for a compiler from \srccm{} to \trgcm{} (Section~\ref{sec:fa-proof}) by
  \item using and defining a cross-language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
  \end{itemize}
\end{itemize}
This paper is accompanied by a technical report~\citep{technical_report} with the elided details and proof.

@@

%\FloatBarrier
% \begin{itemize}
% \item present our \emph{target language} and its operational semantics (excerpts)
% \item mention roughly what components look like
% \end{itemize}
\section{Linear Stack and Return Capabilities}
\label{sec:stktokens-explained}
In this section, we introduce our calling convention \stktokens{} that ensures LSE and WBCF.
We will gradually explain each of the security measures \stktokens{} take and motivate them with the attacks they prevent.

@@

%informally explain how an adversary may try to abuse stack and return caps
While the above adds extra protection, it is not sufficient to enforce WBCF and LSE.
The caller provides the adversary with a stack capability and a return pair that they are supposed to use for the call, but the adversary can store the provided capabilities on the heap in order to use them later.
Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
In both examples our component and some adversarial component have been taking turns calling each other, so the stack now contains four stack frames alternating between ours and theirs.
The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) illustrates how we try to ensure LSE by restricting the stack capability to the unused part before every call to the adversary.
However, restricting the stack capability does not help when we in the first call give access to the part of the stack where our second stack frame will reside as nothing prevents the adversary from duplicating and storing the stack pointer.
Generally speaking, we have no reason to ever trust a stack capability received from a component we do not trust as that stack capability may have been duplicated and stored for later use.
In the figure on the right (Figure~\ref{fig:ret-ptr-abuse}), we have given the adversary two pairs of sealed return capabilities, one in each of the two calls to the adversarial component.
The adversary may have stored the pair of sealed return capabilities from the first call in order to use it in the second call where they are not allowed to use it.
The figure illustrates how the adversarial code uses the return pair from the first call to return from the second call and thus break WBCF.

@@

Register files are related either by a trusted or untrusted relation $\mathcal{R}_\trust(W)$, and they are related when they contain words related in the appropriate value relation.
The register-file relation splits the ownership of $W$ between the registers as linear capability need a spatially owned region; but to prevent aliasing, the they cannot depend on the same spatially owned region.
% memory relation
The memory on \srccm{} consists of a heap $\ms_S$, a call stack $\stk$ and a free stack $\ms_\stk$.
This relates to a memory $\ms_T$ on \trgcm{} in a world $W$ if we can split $\ms_T$ into three parts that correspond with the view \srccm{} provides on the memory.
Each of the three partitions of $\ms_T$ must be related to one of the three parts of \srccm{} memory under the relevant sub-world of $W$.
The heap $\ms_S$ relates to the heap partition of $\ms_T$ in the standard way, i.e.\ the two memories can be partitioned, so they satisfy all the regions of the heap world.
The free stack $\ms_\stk$ also relates to the free stack partition of $\ms_T$ in the standard way but with respect to the free stack world.
%Further, the address $\stkb$ should always be an address of the free stack (the return token always has authority over $\stkb$), so it must be one of the addresses of $\ms_\stk$.
The call stack $\stk$ relates to the call stack partition of $\ms_T$ if it can be split into parts related to each of the stack frames.
Further, as each of the stack frames corresponds to a slot in the return order none of the stack frames should be empty.
It is the physical location of a stack frame that determines its order in the call stack, so the order must be preserved by the partitions of $\ms_T$.

>>>>

%The main contributions of this paper are (1) \stktokens{} a calling convention that uses linear capabilities to provably secure LSE and WBCF, and (2) a novel proof technique that we call fully-abstract overlay semantics which we apply to prove the properties of \stktokens{}.
In this paper, we make the following contributions:
% Presents the first formal model of a capability machine with linear capabilities (Section 2)
% Presents a calling convention called \stktokens{} that provably guarantee local state encapsulation and well-bracketed control flow (Section 3)
% Presents a novel proof technique that we call fully-abstract overlay semantics. (Section 4-5)
\begin{itemize}
\item We present \trgcm{}: A formalization of a simple CHERI-like capability machine with linear capabilities (Section~\ref{sec:cap-mach-w-seal-and-lin}).
\item We present a new calling convention \stktokens{} that provably guarantees LSE and WBCF on \trgcm{} (Section~\ref{sec:stktokens-explained}).
\item We present a new way to formalize these guarantees based on a novel
  technique called \textit{fully-abstract overlay semantics} and we prove LSE
  and WBCF claims. This includes:
  \begin{itemize}
  \item defining \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF (Section~\ref{sec:form-secur-with})
  \item proving full-abstraction for a compiler from \srccm{} to \trgcm{} (Section~\ref{sec:fa-proof}) by
  \item using and defining a cross-language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
  \end{itemize}
\end{itemize}
This paper is accompanied by a technical report~\citep{technical_report} with the elided details and proof.

@@

%\FloatBarrier
% \begin{itemize}
% \item present our \emph{target language} and its operational semantics (excerpts)
% \item mention roughly what components look like
% \end{itemize}
\section{Linear Stack and Return Capabilities}
\label{sec:stktokens-explained}
In this section, we introduce our calling convention \stktokens{} that ensures LSE and WBCF.
We will gradually explain each of the security measures \stktokens{} takes and motivate them with the attacks they prevent.

@@

%informally explain how an adversary may try to abuse stack and return caps
While the above adds extra protection, it is not sufficient to enforce WBCF and LSE.
The caller provides the adversary with a stack capability and a return pair that they are supposed to use for the call, but the adversary can store the provided capabilities on the heap in order to use them later.
Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
In both examples our component and some adversarial component have been taking turns calling each other, so the stack now contains four stack frames alternating between ours and theirs.
The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) illustrates how we try to ensure LSE by restricting the stack capability to the unused part before every call to the adversary.
However, restricting the stack capability does not help when we, in the first call, give access to the part of the stack where our second stack frame will reside as nothing prevents the adversary from duplicating and storing the stack pointer.
Generally speaking, we have no reason to ever trust a stack capability received from a component we do not trust as that stack capability may have been duplicated and stored for later use.
In the figure on the right (Figure~\ref{fig:ret-ptr-abuse}), we have given the adversary two pairs of sealed return capabilities, one in each of the two calls to the adversarial component.
The adversary may have stored the pair of sealed return capabilities from the first call in order to use it in the second call where they are not allowed to use it.
The figure illustrates how the adversarial code uses the return pair from the first call to return from the second call and thus break WBCF.

@@

Register files are related either by a trusted or untrusted relation $\mathcal{R}_\trust(W)$, and they are related when they contain words related in the appropriate value relation.
The register-file relation splits the ownership of $W$ between the registers as linear capability need a spatially owned region; but to prevent aliasing, they cannot depend on the same spatially owned region.
% memory relation
The memory on \srccm{} consists of a heap $\ms_S$, a call stack $\stk$ and a free stack $\ms_\stk$.
This relates to a memory $\ms_T$ on \trgcm{} in a world $W$ if we can split $\ms_T$ into three parts that correspond with the view \srccm{} provides on the memory.
Each of the three partitions of $\ms_T$ must be related to one of the three parts of \srccm{} memory under the relevant sub-world of $W$.
The heap $\ms_S$ relates to the heap partition of $\ms_T$ in the standard way, i.e.\ the two memories can be partitioned, so they satisfy all the regions of the heap world.
The free stack $\ms_\stk$ also relates to the free stack partition of $\ms_T$ in the standard way but with respect to the free stack world.
%Further, the address $\stkb$ should always be an address of the free stack (the return token always has authority over $\stkb$), so it must be one of the addresses of $\ms_\stk$.
The call stack $\stk$ relates to the call stack partition of $\ms_T$ if it can be split into parts related to each of the stack frames.
Further, as each of the stack frames corresponds to a slot in the return order none of the stack frames should be empty.
It is the physical location of a stack frame that determines its order in the call stack, so the order must be preserved by the partitions of $\ms_T$.
